---
title: "R assessment Quiz"
author: "Ivy Fernandes"
date: "3/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R assessment Quiz
What is the minimum input required to produce a single-layer bar plot in R
only need the height vecotr, therefore one numeric vector or matrix
```{r}
height<-(1:30)
barplot(height)
```

data frame vs matrix: A data frame is more general than a matrix, in that different columns can have different modes (numeric, character, factor, etc.). ... Matrix - A matrix is a collection of data elements arranged in a two-dimensional rectangular layout. DataFrame - A dataframe is used for storing data tables.

data frame vs vector: A vector has 1 dimension while a data frame has 2. ... A better question is what is the difference between a data frame and a matrix: a data frame can have different data types for each column, while a matrix is all one data type. Behind the scenes, a data frame is really a list with equal length vectors at each index.


---

jal <- function(x) {y=x^2; return(y)} #If I want to execute jal(hmk), what kind of data type can hmk be without returning an error message? Check all that apply: factor, character, numeric, logical
Factors in R. Conceptually, factors are variables in R which take on a limited number of different values; such variables are often refered to as categorical variables. ... Factors in R are stored as a vector of integer values with a corresponding set of character values to use when the factor is displayed.x
```{r}
jal <- function(x) {y=x^2; return(y)} 
```


What is the minimum input required to produce a single-layer histogram in R

```{r}
h=rnorm(n=30)
View(h)
hist(h)
```

