---
title: 'class 11: structural bioinformatics'
author: "Ivy Fernandes"
date: "2/15/2019"
output: html_document
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##
Getting the data from the PDB website
```{r}
library(readr)
pdbdata <- read.csv("pdbdata.csv", row.names=1)
head(pdbdata)
```

Determine the percentage of structures solved by X-Ray and Electron Microscopy. Also determine what proportion of structures are protein? 

```{r}
totals<-sum(pdbdata$Total)
totals

xraypercentage<-100*(pdbdata$Total[1]/totals)
xraypercentage

empercentage<-100*(pdbdata$Total[3]/totals)
empercentage

proteinpercentage<-100*(sum(pdbdata$Proteins)/totals)
proteinpercentage
```
the other 7% of non-protein things in the data-base are complexed with the proteins

using data pasta
```{r}
data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "NMR", "Electron Microscopy", "Other",
                           "Multi Method", "Total"),
              Proteins = c(124770, 10988, 2057, 250, 127, 138192),
         Nucleic.Acids = c(1993, 1273, 31, 4, 5, 3306),
    ProteinNA.Complex = c(6451, 257, 723, 6, 2, 7439),
                 Other = c(10, 8, 0, 13, 1, 32),
                 Total = c(133224, 12526, 2811, 273, 135, 148969)
)

```



Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB? 

2880 protien structures 
1138 of them are proteases



# Bio3d
```{r}
library(bio3d)
```

read the pdb file
```{r}
pdb<-read.pdb("1hsg")
pdb
```
printing the information from the pdb will tell you the molecules in the pdb and what makes up the structure 

# structure viewing
```{r}
# The 'devtools' package allows us to install development versions
#install.packages("devtools")
# Install the bio3d.view package from bitbucket
#devtools::install_bitbucket("Grantlab/bio3d-view")
# Load the package
library("bio3d.view")
# view the 3D structure
view(pdb, "overview", col="sse")
```

extract the protein only portion of this PDB structure and write it out to a new PDB file

extract the ligan (drug) - going back to section 4 of hands-on
```{r}
# Select all C-alpha atoms (return their indices)
ca.inds <- atom.select(pdb, "calpha")
ca.inds

# Print details of the first few selected atoms
head( pdb$atom[ca.inds$atom, ] )

# And selected xyz coordinates
head( pdb$xyz[, ca.inds$xyz] )
```


trying to select out ligand
```{r}
ligand.inds<-atom.select(pdb,"ligand")
ligand.inds
```

find-grained atom selection 
```{r}
ligand.inds$atom
```
prints out a bunch of numbers, these are the indeces to get at the information within the structure

```{r}
pdb$atom
pdb$atom[ligand.inds$atom,]
```
uses the indeces to get the values 

trim.pdb: Produce a new smaller PDB object, containing a subset of atoms, from a given larger PDB object.

we are aiming to trim down everything so that it doesnt have the water and other molecules but just the ligand

```{r}
ligand.pdb<- trim.pdb(pdb,ligand.inds)
ligand.pdb
```

shows that there are 0 protein atoms and that the only residue is mk1

```{r}
view(ligand.pdb)
```

```{r}
write.pdb(ligand.pdb,file="ligand.pdb")
```


writing the protein file only!
```{r}
protein.inds<-atom.select(pdb,"protein")
protein.inds


pdb$atom[protein.inds$atom,]

protein.pdb<-trim.pdb(pdb, protein.inds)
protein.pdb

view(protein.pdb)
```



```{r}
pdb2<-read.pdb("1HEL")

#Normal Mode Analysis
modes<-nma(pdb2)
plot(modes)
```

