---
title: 'Class 12: Structural Bioinformatics'
author: "Ivy Fernandes"
date: "2/20/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clean up our protein target surface

```{r}
library(bio3d)

pdb.code<-"1hsg" 
#allows for you to only have to change this one area if you want to re-use this for other proteins in the future 

file.name<-get.pdb(pdb.code)

```

need to read the structure into R.  extract the protein only segment of this PDB entry and write out a new PDB format file.  
```{r}
hiv<-read.pdb(file.name)
hiv #to get a summary of what is in the file
```
The summary shows that besides the protein, there is also the ligand (MK1) and water molecules

first extract the protein
```{r}
prot<-trim.pdb(hiv,"protein")
prot
```

```{r}
prot.filename<-paste(pdb.code,"_protein.pdb", sep="") #will print out the protein that you inititally serarched for attatched to this new string, must use the sep="" so no spaces are incorporated
write.pdb(prot,prot.filename) #writes the data to a file
```

now do the same for the ligand
```{r}
ligand<-trim.pdb(hiv,"ligand")
ligand

ligand.filename<-paste(pdb.code,"_ligand.pdb",sep="")
write.pdb(ligand,ligand.filename)
```

## Using autodock tools to setup protein docking

PDB files do not include charges and atom types that describe properties, therefore the files for the proteins and the ligands have to be prepared to include these values along with the atomic coordinates

## Docking indinavir into HIV-1 protease 
```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

Loaded into VMD

Qualitatively, how good are the docks? Is the crystal binding mode reproduced? Is it the
best conformation according to AutoDock Vina?

Quantitatively assess the docking using root mean square distance
```{r}
ori <- read.pdb("1hsg_ligand.pdbqt")
rmsd(ori, res)
```
Quantitatively how good are the docks? Is the crystal binding mode reproduced within 1Å
RMSD for all atoms?

## Exploring the conformational dynamics of protein 
