---
title: "Class 13: Genome informatics and high throughput sequencing"
author: "Ivy Fernandes"
date: "2/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
data<-read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
data
```

What proportion of the Mexican Ancestry in Los Angeles sample population (MXL) are homozygous for the asthma associated SNP (G|G)? 

Want to look only at the second column and look for G|G
```{r}
homozygous_test<-data[,2]=="G|G"
homozygous_test
homozygous_count<-sum(homozygous_test)
homozygous_count
```

other method
```{r}
genotype<-data$Genotype..forward.strand.
table(genotype)
```
counts it out for you


percentage
```{r}
9/nrow(data)*100
```


##Initial RNA-Seq analysis
Now, you want to understand whether the SNP will affect gene expression
HG00109 Sample

```{r}
library(seqinr) 
library(gtools)
chars<-s2c("DDDDCDEDCDDDDBBDDDCC@")
#s2c takes single string and makes each letter an individual vector
phred <-asc(chars)-33
phred
```

## Population scale analysis
```{r}
populationdata<-read.table("rs8067378_ENSG00000172057.6.txt")
head(populationdata)
```
```{r}
table(populationdata$geno) # to find out how many g/g genotypes are
```
want to know of the 121 samples, what the medium expression of a particular gene is
must find out where the columns / rows are g/g

```{r}
gg<-populationdata$geno== "G/G"
gg

```

```{r}
match(populationdata$geno,"G/G")
```

use the logical indeces to get the expression

```{r}
exp_gg<-populationdata[gg, "exp"]
exp_gg
summary(exp_gg)
```


what about for aa?
```{r}
aa<-populationdata$geno== "A/A"
aa
exp_aa<-populationdata[aa, "exp"]
exp_aa
summary(exp_aa)
```


what about for ag?
```{r}

ag<-populationdata$geno== "A/G"
ag
exp_ag<-populationdata[ag, "exp"]
exp_ag
summary(exp_ag)

```


Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?
```{r}
boxplot(exp ~ geno, data=populationdata, notch=TRUE)
```

 

