---
title: "Class 06"
author: "Ivy Fernandes"
date: "January 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## File Reading Again
here we try to use ** read.table()** and friends

utilizing the insert button to insert code chunks -- going to insert {r} code
can put a url into the read.table function 
```{r}
read.table("https://bioboot.github.io/bggn213_S18/class-material/test1.txt", header=TRUE, sep=",")
```
shortcut to get a new code chunk is ctrl+alt+i
```{r}
file1<-"https://bioboot.github.io/bggn213_S18/class-material/test1.txt"
read.csv(file1)

```
```{r}
file2<-"https://bioboot.github.io/bggn213_S18/class-material/test2.txt"
read.table(file2, header=TRUE,sep='$')
```

```{r}
file3<-"https://bioboot.github.io/bggn213_S18/class-material/test3.txt"
data3=read.table(file3)
data3
```

## Now we will start working with functions
Creating the first function
```{r}
add<-function(x,y=1){
  #sum x and y
  x+y
}
```

time to use the **add()** function
```{r}
add(1)
add(1,5) # will replace the y=1 with 5 because you have included an argument
add(c(1,2,3,4)) #will add 1 to all of these numbers within the vector
add(c(1,2,3,4),4)
# add(1,2,2) will not work because there are too many arguments
# add(x=1,y="b") will not work because there is a non-numeric value
```

Create a second function
```{r}
rescale<-function(x){
rng<-range(x)
(x-rng[1]/(rng[2]-rng[1]))
}
```
testing the second function
```{r}
rescale(1:10)
rescale(5:10)
rescale(11:15)
```

Test the function until it fails
```{r}
rescale(c(1,2,NA,3,10)) #missing data point, what WOULD you want your function to do if you have data missing?
```
Testing the function to determine the errors
```{r}
range(c(1,2,NA,3,10))
```
Look at range function to figure out if we can work around giving NA and working with functions that do exist
Identified argument within function that ignores missing values na.rm=FALSE should be changed to TRUE
Edit original function so that it will work

```{r}
rescale2<-function(x){
rng<-range(x,na.rm=TRUE)
(x-rng[1]/(rng[2]-rng[1]))
}
```

Test editted functions
```{r}
rescale2(c(1,2,NA,3,10))
```
Trying to use non-numeric will result in an error
```{r}
# rescale(c(1,2,NA,3,10,"bggn213")) can potentially incorporate warnings that there are non-numerics 
```

How to print lines out in code-- use PRINT FUNCTION
```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 if(na.rm) {
 rng <-range(x, na.rm=na.rm)
 } else {
 rng <-range(x)
 }
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
   print("please don't ever sing again") # how come if plot=TRUE does it decide not to print this, why doesn't it do both
 }
 print("I can see it in ...")
}
```

```{r}
rescale3(1:10)
rescale3(1:10,plot=TRUE)
```
```{r}
# rescale(1:10,plot=FALSE)
```

#Hands-on Worksheet

Using the bio3d package
```{r}
library(bio3d)
```

```{r}
pdb<- read.pdb("1hbs") #accession code 
```

```{r}
pdb
```
copy paste code from document
```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

first and third plots are the same -- looking at the code, the inputs are wrong and s1 is used twice 

```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

